---
title: "Final Project"
format: html
---

### Exploratory Data Analysis

First we can run some exploratory data analysis

multixikkubearutr


```{r}

library(readr)
library(tidyverse)
library(janitor)
library(pheatmap)
library(corrplot)
music_dataset <- read_csv("music_dataset.csv")


ggplot(aes)

```


```{r}
tidy_music_data <- clean_names(music_dataset)

View(tidy_music_data)
```


#Variables

#MAybe create a map of all the variables to see if any are correlated with each other then look closer at pais of varibles


```{r}
predictor_variables <-
  select(tidy_music_data, -c("song","artist", "streams", "daily_streams", "genre", "release_year"))
  
```


#heatmap
```{r}
View(predictor_variables)
pheatmap(predictor_variables, scale = "none")

```



```{r}
cor_matrix <- cor(predictor_variables)
View(cor_matrix)
```

```{r}
pheatmap(cor_matrix, scale = "none")
```
```{r}
corrplot(cor_matrix,method="pie")
```
```{r}
corrplot(cor_matrix,method="shade")

```

```{r}
high_correlations <- cor_matrix[abs(cor_matrix) > 0.5]
View(high)
```
```{r}
cor_matrix <- cor(predictor_variables, use = "complete.obs")
corrplot(cor_matrix, method = "color")
```


Our variables seem to not have any correlations with each other
# Danceability + Energy

Since these two variables might be simlar, we might want to first investigate these variables to capture any multicolinnarity. These vairables seem to be numeric and continuos, 










